# Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```



```{r out.width="800px"}
a <- movies |> distinct(studio,year,title) |>
  group_by(year,studio) |>
  mutate(studio = if_else(studio %in% (movies |>
                                      distinct(studio, gross_CPI_adjusted) |>
                                      group_by(studio) |>
                                      summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
                                      arrange(desc(gross_CPI_adjusted)) |>
                                      filter(!is.na(studio)) |>
                                      slice(1:7) |>
                                      select(studio) |> unlist()), studio,"Other"))|>
  filter(!is.na(studio)) |>
  ggplot(aes(x=year, fill=factor(studio)))+
  geom_bar(position = "fill")+
  labs(title = "Other Studios Are Starting To Take More Market Share\nIn Terms of Output But Not Gross",
       subtitle = "Despite the relatively few movies produced by Buena Vista, it maintains spot of top\nGrossing Studios",
       y="Proportion of Movies Produced")+
  theme_minimal()+
  theme(legend.position = "none")

b <- movies |> distinct(studio,year,title,.keep_all = T) |>
  group_by(year,studio) |>
  mutate(studio = if_else(studio %in% (movies |>
                                        distinct(studio, year,gross_CPI_adjusted) |>
                                        group_by(studio) |>
                                        summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
                                        arrange(desc(gross_CPI_adjusted)) |>
                                        filter(!is.na(studio)) |>
                                        dplyr::slice(1:7) |>
                                        select(studio) |> unlist()), studio,"Other")) |>
  filter(!is.na(studio)) |>
  ggplot(aes(x=year, y=gross_CPI_adjusted,fill=factor(studio)))+
  geom_bar(stat = "identity",position = "fill")+
  labs(y="Proportion of Movies Gross", legend = "studio")+
  theme_minimal()+
  theme(legend.position = "bottom", legend.key.size = unit(0.5, 'cm'))

(a | b)

```


Now we break the model into impacts of Genre, Run time, .....nkdfjsdkfjlsddöf



* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

Here we analyse the variables we believe have an influence on movie success

## Genre - +geom_smooth()

* Genre

* Run-time (Attention span)

* Budget
Exclude all movies with budgets that are not US$.

* Certificate

* Studio

* Director

* Actor