---
title: "Mohamed"
format: html
editor: visual
---

## Quarto

```{r}
library(ggplot2)
library(dplyr)

movies |> distinct(dob_stars,stars) |>
  mutate(dob_stars = year(dob_stars)) |>
  group_by(dob_stars) |>
  summarise(count = n()) |>
  filter(!is.na(dob_stars)) |>
  ggplot(aes(x=dob_stars,y=count)) +
  geom_point()


movies |> distinct(dob_directors,directors) |>
  mutate(dob_directors = year(dob_directors)) |>
  group_by(dob_directors) |>
  summarise(count = n()) |>
  filter(!is.na(dob_directors)) |>
  ggplot(aes(x=dob_directors,y=count)) +
  geom_point()

```

```{r}
movies |> distinct(dob_stars,stars) |>
  mutate(dob_stars = month(dob_stars)) |>
  group_by(dob_stars) |>
  summarise(count = n()) |>
  filter(!is.na(dob_stars)) |>
  ggplot(aes(x=dob_stars,y=count)) +
  geom_bar(stat = "identity")


movies |> distinct(dob_stars,stars) |>
  mutate(dob_year = year(dob_stars),
         dob_month = month(dob_stars)) |>
  group_by(dob_year,dob_month) |>
  summarise(count = n()) |>
  ggplot(aes(x=dob_year, y=count, col = factor(dob_month)))+
  scale_y_continuous(limits = c(0,30))+
  coord_cartesian(xlim = c(1900, 2000))+
  geom_line()

```

```{r}
# actors with highest gross
movies |> distinct(stars,gross,dob_stars,gross_CPI_adjusted) |>
  group_by(stars,dob_stars) |>
  reframe(gross = sum(gross,na.rm=T),
            gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = T)) |>
  filter(stars!="") |>
  arrange(desc(gross_CPI_adjusted))


```
# Studio
```{r}

# Crazy Netflix
movies |> distinct(title,studio,year) |>
  group_by(studio,year) |>
  reframe(count = sum(!is.na(title))) |>
  filter(!(studio==""|studio=="Studio"|is.na(studio)),count>10) |>
  ggplot(aes(x=year,y=count,col=factor(studio)))+
  geom_line()


movies |>
  distinct(title, studio, year) |>
  filter(studio %in% (
    movies |>
    distinct(studio, gross_CPI_adjusted) |>
    group_by(studio) |>
    summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
    arrange(desc(gross_CPI_adjusted)) |>
    filter(!is.na(studio)) |>
    slice(1:15) |>
    select(studio) |> unlist()
  ))

movies |>
    distinct(studio, gross_CPI_adjusted) |>
    group_by(studio) |>
    summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
    arrange(desc(gross_CPI_adjusted)) |>
    filter(!is.na(studio)) |>
    slice(1:5) |>
    select(studio) |> left_join(movies |> distinct(title,studio,year), join_by(studio)) |>
  group_by(studio,year) |>
  reframe(count = n()) |>
  ggplot(aes(x=year,y=count,col=studio))+
    geom_line()

#check avg gross per year maybe some studios are improving
# To check studio's avg ratings per year

# Harder to please the audience
movies |> distinct(year,studio,rating)|>
  group_by(studio,year)|>
  reframe(mean_rating = mean(rating)) |>
  right_join(movies |>
    distinct(studio, gross_CPI_adjusted) |>
    group_by(studio) |>
    summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
    arrange(desc(gross_CPI_adjusted)) |>
    filter(!is.na(studio)) |>
    dplyr::slice(1:20) |>
    select(studio),join_by(studio)) |>
  ggplot(aes(x=year,y=mean_rating,col=factor(studio)))+
  geom_line()


```

```{r}
movies |> distinct(title,year,studio,gross_CPI_adjusted) |>
  group_by(studio) |>
  summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted,na.rm = T)) |>
  filter(!is.na(studio)) |> ungroup() |>
  left_join(movies |> distinct(title,rating,studio) |>
    arrange(desc(rating)) |>
    group_by(studio) |>
    summarise(best_title = first(title)),join_by(studio)) |>
  arrange(desc(gross_CPI_adjusted)) |>
  slice(1:20) |> 
  ggplot(aes(x=fct_reorder(studio,gross_CPI_adjusted),y=gross_CPI_adjusted))+
  geom_bar(stat = "identity") +
  geom_text(aes(label = best_title, vjust = 0),angle = 0, hjust = 0,size=4)+
  theme_minimal()+
  scale_y_continuous(labels = scales::comma, limits = c(0,120e9))+
  labs(title = "")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  coord_flip()


movies |> distinct(title,year,studio,gross_CPI_adjusted) |>
  group_by(studio) |>
  summarise(gross_CPI_adjusted = mean(gross_CPI_adjusted,na.rm = T)) |>
  filter(!is.na(studio)) |> ungroup() |>
  left_join(movies |> distinct(title,rating,studio) |>
    arrange(desc(rating)) |>
    group_by(studio) |>
    summarise(best_title = first(title)),join_by(studio)) |>
  arrange(desc(gross_CPI_adjusted)) |>
  slice(1:20) |> 
  ggplot(aes(x=fct_reorder(studio,gross_CPI_adjusted),y=gross_CPI_adjusted))+
  geom_bar(stat = "identity") +
  geom_text(aes(label = best_title, vjust = 0),angle = 0, hjust = 0,size=4)+
  theme_minimal()+
  #scale_y_continuous(labels = scales::comma, limits = c(0,120e9))+
  labs(title = "")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  coord_flip()

movies |> distinct(title,year,studio,gross_CPI_adjusted,budget_CPI_adjusted) |>
  group_by(studio) |>
  summarise(profitability = sum((gross_CPI_adjusted-budget_CPI_adjusted)/budget_CPI_adjusted,na.rm = T)) |>
  filter(!is.na(studio)) |> ungroup() |>
  #left_join(movies |> distinct(title,rating,studio) |>
  #  arrange(desc(rating)) |>
  #  group_by(studio) |>
  #  summarise(best_title = first(title)),join_by(studio)) |>
  arrange(desc(profitability)) |>
  dplyr::slice(1:20) |> 
  ggplot(aes(x=fct_reorder(studio,profitability),y=profitability))+
  geom_bar(stat = "identity") +
  #geom_text(aes(label = best_title, vjust = 0),angle = 0, hjust = 0,size=4)+
  theme_minimal()+
  #scale_y_continuous(labels = scales::comma, limits = c(0,120e9))+
  labs(title = "")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))+
  coord_flip()

```



```{r}
a <- movies |> distinct(studio,year,title) |>
  group_by(year,studio) |>
  mutate(studio = if_else(studio %in% (movies |>
                                      distinct(studio, gross_CPI_adjusted) |>
                                      group_by(studio) |>
                                      summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
                                      arrange(desc(gross_CPI_adjusted)) |>
                                      filter(!is.na(studio)) |>
                                      slice(1:7) |>
                                      select(studio) |> unlist()), studio,"Other"))|>
  ggplot(aes(x=year, fill=factor(studio)))+
  geom_bar(position = "fill")+
  labs(title = "Other Studios Are Starting To Take More Market Share\nIn Terms of Output But Not Gross",
       subtitle = "Despite the relatively few movies produced by Buena Vista, it maintains spot of top\nGrossing Studios",
       y="Proportion of Movies Produced")+
  theme_minimal()+
  theme(legend.position = "none")

b <- movies |> distinct(studio,year,title) |>
  group_by(year,studio) |>
  mutate(studio = if_else(studio %in% (movies |>
                                        distinct(studio, year,gross_CPI_adjusted) |>
                                        group_by(studio) |>
                                        summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
                                        arrange(desc(gross_CPI_adjusted)) |>
                                        filter(!is.na(studio)) |>
                                        dplyr::slice(1:7) |>
                                        select(studio) |> unlist()), studio,"Other"))|>
  
  ggplot(aes(x=year, y=gross_CPI_adjusted,fill=factor(studio)))+
  geom_bar(stat = "identity",position = "fill")+
  labs(y="Proportion of Movies Gross")+
  theme_minimal()

(a | b)

```




Analysis on the past 5-10 years most movies produced simply a filter!
```{r}

b <- movies |> distinct(studio,year,title,.keep_all = T) |>
  group_by(year,studio) |>
  #summarise(n()) |>
  mutate(studio = if_else(studio %in% (movies |>
                                        distinct(studio, year,gross_CPI_adjusted) |>
                                        group_by(studio) |>
                                        summarise(gross_CPI_adjusted = sum(gross_CPI_adjusted, na.rm = TRUE)) |>
                                        arrange(desc(gross_CPI_adjusted)) |>
                                        filter(!is.na(studio)) |>
                                        dplyr::slice(1:7) |>
                                        select(studio) |> unlist()), studio,"Other"))|>
  ggplot(aes(x=year, y=gross_CPI_adjusted,fill=factor(studio)))+
  geom_bar(stat = "identity",position = "fill")+
  labs(title = "It Seems that Other Studios Are Starting To Take More Market Share In Terms of\nNumber of Movies Produced",
       subtitle = "Despite the few movies produced by Buena Vista, it maintains spot of top\nGrossing Studios",
       y="Proportion of Movies Gross")+
  theme_minimal()

```

## A Naiive Approah?
```{r}

model_set <- movies |> 
  distinct(title,year,stars,directors,dob_directors,dob_stars,certificate_adjusted,run_time,genre,gross_CPI_adjusted,budget_CPI_adjusted,
           studio) |>
  mutate(stars_age = year-year(dob_stars),
         directors_age = year-year(dob_directors))|>
  filter(stars_age>0&stars_age<100,directors_age>0&directors_age<100) |>
  group_by(title,genre,dob_directors) |>
  mutate(stars_avg_age = mean(stars_age,na.rm = T),
         directors_avg_age = year-year(dob_directors),
         nr_stars = n()) |>
  ungroup() |>
  group_by(title,genre,dob_stars) |>
  mutate(nr_directors = n()) |>
  ungroup() |>
  select(-dob_stars,-stars_age,-dob_directors,-directors_age) |>
  distinct() |>
  filter(!is.na(gross_CPI_adjusted-budget_CPI_adjusted)) |>
  mutate(is_successful = if_else(gross_CPI_adjusted-budget_CPI_adjusted>0,1,0) |> factor(),
         certificate_adjusted = if_else(is.na(certificate_adjusted),0,certificate_adjusted),
         studio = factor(studio),
         genre = factor(genre)) |>
  left_join(movies |> 
              distinct(title,studio) |> 
              group_by(studio) |> 
              summarise(nr_movies_p_studio = n()),join_by(studio)) |>
  left_join(movies |> 
              distinct(stars,title) |> 
              group_by(stars) |> 
              summarise(nr_movies_p_stars = n()),join_by(stars)) |>
  left_join(movies |> 
              distinct(directors,title) |> 
              group_by(directors) |> 
              summarise(nr_movies_p_directors = n()),join_by(directors))|>
  select(-gross_CPI_adjusted,-year,-title,-stars,-directors)



model_set <- movies |> 
  distinct(title,year,stars,directors,dob_directors,dob_stars,certificate_adjusted,run_time,gross_CPI_adjusted,budget_CPI_adjusted,
           studio) |>
  mutate(stars_age = year-year(dob_stars),
         directors_age = year-year(dob_directors))|>
  filter(stars_age>0&stars_age<100,directors_age>0&directors_age<100) |>
  group_by(title,dob_directors) |>
  mutate(stars_avg_age = mean(stars_age,na.rm = T),
         directors_avg_age = year-year(dob_directors),
         nr_stars = n()) |>
  ungroup() |>
  group_by(title,dob_stars) |>
  mutate(nr_directors = n()) |>
  ungroup() |>
  select(-dob_stars,-stars_age,-dob_directors,-directors_age) |>
  distinct() |>
  filter(!is.na(gross_CPI_adjusted-budget_CPI_adjusted)) |>
  mutate(is_successful = if_else(gross_CPI_adjusted-budget_CPI_adjusted>0,1,0) |> factor(),
         certificate_adjusted = if_else(is.na(certificate_adjusted),0,certificate_adjusted),
         studio = factor(studio)) |>
  left_join(movies |> 
              distinct(title,studio) |> 
              group_by(studio) |> 
              summarise(nr_movies_p_studio = n()),join_by(studio)) |>
  left_join(movies |> 
              distinct(stars,title) |> 
              group_by(stars) |> 
              summarise(nr_movies_p_stars = n()),join_by(stars)) |>
  left_join(movies |> 
              distinct(directors,title) |> 
              group_by(directors) |> 
              summarise(nr_movies_p_directors = n()),join_by(directors))|>
  #left_join(movies|>distinct(title,year,directors) |>
  #            group_by(directors) |>
  #            mutate(rank = 1+map_int(year,~sum(.x>year,na.rm=T)))|> select(directors,rank,year,title),join_by(directors,year,title)) |>
  select(-gross_CPI_adjusted,-year,-title,-stars,-directors,nr_movies_p_stars)

model_set <- model_set[complete.cases(model_set),]


model_glm <- glm(is_successful ~ ., data = model_set, family = binomial)
broom::glance(model_glm)
broom::tidy(model_glm)

library("ROCR")
library("caret")


## cross validated model
set.seed(123)  # Set a random seed for reproducibility
folds <- createFolds(model_set$is_successful, k = 10)

ctrl <- trainControl(method = "cv", number = 10, index = folds, savePredictions = "final")


model_cross_validated <- train(is_successful ~ ., data = model_set, method = "glm", trControl = ctrl, family = binomial)
model_cross_validated #aic 1783 + accuracy 85% 10fold cross validated 10 order of magnitude improvement to just glm



## Random forest
library(randomForest)

set.seed(123)  # For reproducibility
sample_indices <- sample(nrow(model_set), nrow(model_set) * 0.7)
train_data <- model_set[sample_indices, ]
test_data <- model_set[-sample_indices, ]

response <- "is_successful"
predictors <- c("run_time","certificate_adjusted","budget_CPI_adjusted","directors_avg_age",
                "stars_avg_age", "nr_stars","nr_movies_p_studio","nr_movies_p_directors")

rf_model <- randomForest(
  x = train_data[, predictors],
  y = train_data$is_successful,
  ntree = 100,  # Number of trees in the forest
  mtry = sqrt(length(predictors)),  # Number of features to consider at each split
  importance = TRUE  # Calculate variable importance
)

predictions <- predict(rf_model, newdata = test_data)

# Calculate accuracy
accuracy <- mean(predictions == test_data$is_successful)
cat("Accuracy:", accuracy, "\n")




# RF results:
model_set |> ggplot(aes(x=factor(certificate_adjusted),fill=factor(is_successful)))+geom_bar(position="dodge")


model_set |> ggplot(aes(x=factor(is_successful),y=run_time))+geom_boxplot()

```
So far we considered movies success as being assessed mostly in relation to gross and rating. One way to view success however is from a profitability stand point. Investors and Studios may argue that for a movie to be successful it has to generate profit. For this particular reason we decided to build a model that aims for predicting successful. We consider a movie that has not generated a loss to be "successful". The model was built using random forest and a series of feature engineered variables and it has revealed that the features that explain successfulness best are in descending order:
-
-
-
-
-
-
